Methods for Large Scale Learning to Rank

### 2014-11-19 report/Chapters/LTRMethods.tex
smoothrank beter

where $p_i$ is a probability over documents $i \in \{1,...,n\}$ given the outputs $f_1,...,f_n$ of the evaluation function on the document relevance labels, with $\sigma$ being a smoothing parameter. Chapelle et al. \cite{Chapelle2010} apply this softmax function to the \ac{NDCG} formula and hereby introduce a soft version of indicator variable $h_{i,j}$ through the computation shown in Equation \ref{eq:soft_NDCG}.

@@

It can be shown that the derivative of the smoothed \ac{NDCG} version shown in Equation \ref{eq:soft_NDCG} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(n^2)$, which enable fast gradient descent optimisation. The optimisation step in SmoothRank uses the nonlinear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}, which is a type of gradient descent method. This optimisation method is prone to local optima, which is alleviated by adding a pre-calculated, better than naive, starting point and by adding a regularisation term.\\

@@

The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_NDCG} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function with optima that substantially differ from the true optimal rankings. To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by two at each iteration. Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.\\

>>>>

where $p_i$ is a probability over documents $i \in \{1,...,n\}$ given the outputs $f_1,...,f_n$ of the ranking function on the documents, with $\sigma$ being a smoothing parameter. Chapelle et al. \cite{Chapelle2010} apply this softmax function to the \ac{NDCG} formula and hereby introduce a soft version $h_{i,j}$ of the indicator variable $\mathbb{1}_{i,j}$ ("is document $i$ ranked at the $j$-th position?") through the computation shown in Equation \ref{eq:soft_NDCG}.

@@

It can be shown that the derivative of the smoothed \ac{NDCG} version shown in Equation \ref{eq:soft_NDCG} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(n^2)$, which enable fast gradient descent optimisation. The optimisation step in SmoothRank uses the non-linear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}. This optimisation method is prone to local optima, which is alleviated by adding a pre-calculated, better than naive, starting point and by adding a regularisation term.\\

@@

The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_NDCG} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function with optima that substantially differ from the true optimal rankings. To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing procedure where optimisation starts with a large $\sigma$ in the first iteration and iteratively reduces it by dividing it by two at each iteration. Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.\\
